%% ------------------------------ Abstract ---------------------------------- %%
\begin{abstract}
	
	%In this work, two new reduced-order models (ROMs) for 1D thermal radiative transfer problems are presented. These ROMs are based on the multilevel nonlinear iterative-projective methodology and the data-driven reduced-order modeling methodology known as the proper orthogonal decomposition. The first ROM avoids solving the radiative transfer equation and only makes use of low-order transport systems. The second ROM avoids solving the radiative transfer equation and multigroup low-order transport system, only using an effective grey problem to find a solution. The numerical results show that the first ROM generates more accurate solutions than other well-known ROMs such as $P_1$ and diffusion. This ROM is also parameterized with respect to temperature of incoming radiation on the problem boundary and this is shown to produce similarly accurate results. The second ROM is also shown to have good accuracy, but is demonstrated to have certain limitations that motivate the need for further development.
	
	Thermal radiative transfer (TRT) is a major piece in various multiphysical phenomena which are driven by interaction between photons and matter, such as radiative hydrodynamics. The dimensionality of TRT problems is determined by the radiative transfer (RT) equation. In this work, we study a new approach for developing physics-based RT reduced-order models. We apply an efficient method for solving coupled multiphysics equations that enables one to reduce the dimensionality of the RT problem and combine it  with a decomposition-based approach for model order reduction. We develop two reduced-order models (ROMs) for TRT problems formulated by means of the multilevel nonlinear projective-iterative (MLNPI) methodology and the data-driven reduced-order modeling methodology known as the proper orthogonal decomposition (POD). The proposed multigroup TRT ROM applies  the multilevel quasidiffusion (QD) method with POD of the group QD (Eddington) factors  that carry essential information about the RT high-order solution. The grey TRT ROM uses the effective grey low-order QD equations and POD of the group QD factors and group radiation energy densities.
	
	The obtained numerical results demonstrate that the multigroup TRT ROM with a data set generated by means of a rather low-rank representation of QD factors in space and time sufficiently accurately approximates the solution of TRT problems. As the rank of the approximation is increased, the accuracy of the model gradually improves. The low-rank version of this QD factor data set can be used as a basis for creating efficient ROMs for multiphysics simulations of the evolution of temperature and radiation energy waves. The analysis also  showed that this multigroup TRT  ROM has potential in parametric model reduction for TRT problems. The grey TRT ROM sufficiently accurately  approximates  the solution of the considered TRT problem with low-rank POD of data. The analysis also revealed some limitations of this ROM that motivate the need for further research and development.
	
	
\end{abstract}

%% -------------------------------- Title page ------------------------------ %%
\maketitlepage

%% -------------------------------- Dedication ------------------------------ %%
\begin{dedication}
 \centering To Mary and Helen
\end{dedication}

%% -------------------------------- Biography ------------------------------- %%
%\begin{biography}
%The author was born in a small town \ldots
%\end{biography}

%% ----------------------------- Acknowledgements --------------------------- %%
\begin{acknowledgements}

	I would like to thank my advisor, Dr. Dmitriy Anistratov for his guidance and time spent furthering my understandings of this field. Thank you also to my family, for their support and patience.

\end{acknowledgements}


\thesistableofcontents

\thesislistoftables

\thesislistoffigures
